Neutral ISM, Lyman-Alpha and Lyman-continuum in nearby starburst Haro 11 \footnote{Based on observations with HST-COS, program GO 13017, PI Heckman}

### 2017-01-23 ToSubmit/RiveraThorsen.tex
First revision submitted
Incorporated comments from Göran and Matthew plus referee report, see
details in the newly added coverletter.* files.

\begin{abstract}
Star forming galaxies are believed to be a major source of ionizing
radiation responsible for reionizing the early Universe. Direct
observations of escaping ionizing radiation have however been few and
with escape fractions far below what would be necessary to account for
the ionization history of the Universe. In the local Universe, only a
few handfuls of emitters have been observed with typical escape
fractions of a few percent. It seems the mechanisms regulating this
escape need to be strongly evolving with redshift, if galaxies are to be
held responsible for reionization. Gas mass, star formation feedback and
thus star formation activity are among the main suspect parameters,
known to both regulate neutral gas coverage and evolve with cosmic time.
In this paper, we present a reanalysis of far-UV HST-COS spectra of the
first detected Lyman-continuum leaker in the local Universe, Haro 11, in
which we examine the connections between Lyman-continuum leakage and
Lyman-$\alpha$ line shape and feedback-influenced neutral ISM properties
like kinematics, geometry and column density. We discuss the two
extremes of an optically thin, density bounded ISM and a riddled,
optically thick, ionization bounded ISM and how Haro 11 fits into the
theoretical predictions made from these models. We find that the most
likely ISM model for Haro 11 knot C is one of clumpy neutral medium
embedded in a highly ionized inter-clump medium with a combined covering
fraction of unity and a residual neutral gas column density in the
ionized medium of
$5 \times 10^{13} \leq N_{\rm HI} \leq 4 \times 10^{15}$ cm⁻²; high
enough to be optically thick to Lyman-$\alpha$, but low enough to be at
least partly transparent to Lyman-continuum and undetected in
\ion{Si}{2}.
\end{abstract}

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used archival imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009}; \citet{Hayes2009}; since the absorption lines
treated here have resolution determined by the properties of their
surrounding continuum. We then extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We modelled
vignetting in the COS aperture by multiplying each pixel in the circular
UV-continuum image with an interpolation of the values given in the
throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the aperture
cutouts were collapsed along the cross-dispersion direction, and the
FWHM of the flux distribution along the dispersion direction is reported
as the effective resolution of this observation. This is converted from
arc seconds to COS resolution elements by the factor 0.171 arcsec/resel
reported in the COS instrument handbook \citep{CosHandbook}

\begin{figure}
\centering
\includegraphics[width=1.000\hsize]{./EffResol.pdf}
\caption{Effective resolution estimate. \textbf{Upper left}: Haro 11 in
UV-continuum, data from \citet{Ostlin2009}; \citet{Hayes2009}. Data is
shown on a square root scale and cut levels set for best detail. Inset
in blue tinting is the COS aperture covering the region surrounding knot
B in the terminology of \citet{Vader1993}. N is up, E is left.
\textbf{Upper right}: The extracted, throughput-corrected in-aperture
image, rotated so horizontal corresponds to the dispersion direction of
the spectrograph. \textbf{Lower panel}: The collapsed flux profile in
the aperture, with the shaded region showing the width of the effective
resolution element, which is also given in arc seconds and resolution
elements/bins.}\label{fig:resol}
\end{figure}

The procedure is illustrated in Fig.~\ref{fig:resol}. Here, the upper
left panel shows Haro 11 in UV-continuum using HST imaging data from
\citet{Ostlin2009}; \citet{Hayes2009}, with the COS aperture coverage
from this observation overlaid in blue tinting. The upper right panel
shows the galaxy as the COS saw it, with the region covered by the
aperture cut out, rotated and vignetting-corrected as described above.
Finally, the lower panel shows the flux profile in the aperture,
collapsed along the cross dispersion direction. The FWHM of this
distribution has been adopted as the effective resolution for this
observation. We find $R_{\text{eff}} = 0.3 \arcsec = 1.75$ resolution
elements, which at the wavelength of observed Ly$\alpha$ corresponds to
31 km s⁻¹.

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{./HISLISProfiles.pdf}
\caption{The \ion{Si}{2} and other LIS (\textbf{upper}) and \ion{Si}{4}
(\textbf{lower}) profiles included in this study. Dashes denote data
points masked out due to foreground contamination. Note the
high-velocity component in both \ion{Si}{4} lines, extending out to
$\sim 850$ km s⁻¹.}\label{fig:SingleLines}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.800\hsize]{./AOD-details-example.pdf}
\caption{Illustration of the AOD computations. \textbf{Upper panel}: the
three \ion{Si}{2} lines in this analysis shown together, with the
zero-velocity bin marked with a vertical red line. \textbf{Lower panel}:
From this bin, the three measured relative intensities are plotted
against $f\lambda$ as black dots, with vertical error bars showing the
error spectrum from the COS pipeline. Also shown is the best-fit
function $I/I0 (f \lambda)$ as defined above. Also shown for
illustration are the two extreme examples of a very high column density
and unchanged covering fraction, and of a very high covering fraction
but significantly lower column density.}\label{fig:AOD}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{./LyACoverfracs.pdf}
\caption{\textbf{Upper panel}: Ly$\alpha$ profile of Haro 11 C, in
approximate units of the surrounding continuum level. Full line is the
measured values smoothed by a 5 px. flat kernel; surrounding shading
encloses the $\pm 1 \sigma$ error band (mostly comparable in width to
the line width). \textbf{Middle panel}: Black steps show the averaged,
LIS line profile, smoothed by a 5px kernel. Surrounding gray shading
denotes the $\pm 1 \sigma$ confidence band. \textbf{Lower panel}: Same
as middle panel, but for the \ion{Si}{4} transitions.
}\label{fig:HisLisLya}
\end{figure}

The picture is what we would expect from a Lyman Continuum leaker: weak
neutral and strong ionized absorption profiles, the latter almost
everywhere consistent with full coverage, reveal a highly ionized medium
in front of the central cluster. This is also in good agreement with the
findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the
\ion{Si}{2} and \ion{Si}{4} line profiles, indicating that they likely
represent two different phases in the same higher density regions. These
regions will likely be ionized on the side facing the central cluster,
being photoionized by its hot, massive stars. This also implies that the
nebular emission outflow found in {[}Rivera-Thorsen et al, submitted{]}
at least partially traces the same gas as the \ion{Si}{4} absorption
here and, by extension, also the neutral medium. Looking at
fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found
at around $v \approx -50$ km s⁻¹, where also a component is found in
{[}Rivera-Thorsen et al., submitted{]}. The outflow velocity
$v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully consistent with the
findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium
resonance absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an
overall, weak \emph{redshift} of $v = 32$ km s⁻¹. From
e.g.~fig.~\ref{fig:HisLisLya}, it is evident that while neutral gas
\emph{is} present at these velocities, the velocity is at odds with the
integrated velocity of $v = -149$ km s⁻¹ found in this work. We note
that \ion{Na}{1} has a very low ionization potential of only $\sim 5$
eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.
This is in agreement with the finding of \citet{Sandberg2013} that NaD,
despite being a strong transition, shows absorption of only $\sim 95\%$
of continuum level and is mostly found in small, optically thick clouds.
This suggests that the NaD is tracing only the densest and/or dustiest
regions which are more slowly accelerated by star formation feedback
than the surrounding, more dilute medium. In fig.~\ref{fig:apert}, dusty
regions in the aperture are apparent E and W of knot C, and in the
better resolution of fig. 1 in \citet{Adamo2010}, these regions seem
like they might be connected by a narrow dust lane partially covering
the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of \ion{H}{1} around line center. We find a
Lyman $\alpha$ peak velocity of
$v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$ km s⁻¹ w.r.t the systemic
velocity found by \citet{Sandberg2013}, derived from nebular emission
lines in the region around knot C. This velocity, as it is also
discussed in \citet{Verhamme2015}, is just consistent with their
theoretical predictions for a density-bounded, low-column density
system, albeit on the upper limit of their allowed range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in the same way is in
{[}Puschnig et al., submitted to MNRAS{]} as follows. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$ cm⁻²
for just one velocity bin. Since HI gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario.

Furthermore; while Ly$\alpha$ radiative transfer is dominated by gas of
$v \sim v_0$, Lyman continuum is sensitive to \ion{H}{1} at \emph{all}
velocities. The total column density of \ion{Si}{2} depends on the
configuration of the neutral clouds. Assuming that the area covered by
gas in each velocity bin is completely randomly located, a lower limit
to the total column density is:
%
\begin{equation}
\label{NHImean}
N_{\rm Si II}^{\rm mean} = \sum_i N_{\text{HI}, i} f_{C, i}
\end{equation}
%
 Summing this over $-450$ km s⁻¹ $< v < 150$ km s⁻¹, the velocity range
over which the column densities can be reasonably well determined (and
removing the unphysically high value in the bin at $v \approx 222$ km
s⁻¹), yields a lower limit of $\log N_{\rm Si II} = 13.52 \pm 0.15$ and
$\log N_{\rm H I} 19.01 \pm 0.17$, corresponding to $\approx 64$ optical
depths in Lyman Continuum, strongly incompatible with an optically thin,
density-bounded scenario.

However, the \emph{pure} riddled ionization bounded scenario is easily
ruled out since the Ly$\alpha$ profile does not have any appreciable
emission component at zero velocity. We therefore expect a residual
neutral fraction to remain in the ionized phase; a fraction which has a
column density high enough to block Ly$\alpha$ efficiently at line
center, but low enough to be at least part transparent to Lyman
Continuum radiation and undetected in \ion{Si}{2}.

We can estimate the lowest detectable $N_{\rm Si II}$ by noting that the
relative errors for the \ion{Si}{2} lines in our spectrum are
$\sim 0.05$. Assuming a covering fraction of unity for the dilute
neutral component, and adopting the oscillator strength of the strongest
of the \ion{Si}{2} transitions, $f\lambda_{1260 \AA} = 1486.8$, we find
in the limit that $I/I_0 = e^{-\tau} = 0.95$, and eq.~\ref{eq:tau}
becomes:
%
\begin{equation}
\label{eq:silim}
N_{\rm Si II}^{\rm min} = -\log_{e}(0.95) \frac{3.768 \times 10^{14}}{f\lambda} 
    = 10^{10.1} \text{cm}^{-2},
\end{equation}
%
 which with the adopted metallicity for Haro 11 C corresponds to a
minimum Hydrogen column density of
$N_{\rm HI}^{\rm min} \sim 4.0 \times 10^{15}$ cm⁻². This leaves around
2 orders of magnitude in $N_{\rm H I}$, in which the gas is not detected
in \ion{Si}{2} and is optically thick to Ly$\alpha$ while translucent to
Lyman Continuum. If we require the gas to be detected in at least two of
the lines included in this analysis, the limiting hydrogen column
density becomes $N_{\rm HI}^{\rm min} \sim 4.9 \times 10^{16}$ cm⁻²,
adding another order of magnitude to the allowed range, but we adopt the
lower value as a conservative estimate.

The existence of a diffuse neutral component being present in the
ionized medium seems consistent with what is found in the Green Pea
galaxies of \citep{Henry2015} - these galaxies have sometimes very low
LIS absorption depths, which is usually indicative of low covering
fractions and high porosity of the neutral medium - and yet they find
covering fractions of near unity of \ion{H}{1} from Lyman $\beta$
absorption, indicating column densities of $N \gtrsim 10^{16}$. Like in
the case of Haro 11, this indicates that a non-negligible neutral
component must be present in the ionized phase, although it may very
well be of such low densities and metallicity that metal absorption from
this medium is undetectable.

The column densities we derive for \ion{Si}{2} in this work are
generally between $12.0 \lesssim \log N \lesssim 12.5$, column densities
around which the transitions involved become optically thick: For
$\lambda \lambda 1260, 1526 \text{ and } 1304$, $\tau$ becomes 1 at
$\log N \sim 11.3, 12.2$ and $12.5$, respectively. Off the regions of
strongest absorption, and in particular at $v_0$, only $\lambda 1260$
seems to be optically thick, and it seems the column densities arrived
at can be trusted. But at line center, around $v\sim v_{int}$, the
column densities found are so close to the limit that they are most
likely to be interpreted as lower limits. There might also be
systematics in the determination of the continuum around the lines which
may lead to the lines at $\lambda \lambda 1304, 1526$ to be falsely seen
as shallower than $\lambda 1260$, meaning that all computed column
densities are really lower limits rather than actual values. The
confidence intervals in fig. \ref{fig:WithColDens} do \emph{not} reflect
this possible saturation, but only the formal errors from the best
approximation to the residual intensity, they do not include
systematics.

@@

The exact value of the metallicity is however uncertain. The values
found from nebular recombination lines by \citet{Guseva2012} are
measured mainly in the central \ion{H}{2} regions around the clusters;
and differ by 0.2 dex between knot B and C. The neutral, outflowing gas
could be mixed, or have an unseen LOS distance component larger than the
knot separation, drawing into question which is the better value to
assume for this gas. We base our conclusions on the value found for knot
C, but note that using the metallicity of $12 + \log (O/H) = 8.3$ found
for knot B by \citet{Guseva2012}, the inferred \ion{H}{1} column
densities are $3.9 \pm 0.6 \times 10^{17}$ cm\textsuperscript{-2}.
However, the question of the exact metallicity of the outflowing gas is
complicated. Gas closer to the star forming regions is expected to be
more strongly enriched than gas further away, which would imply that the
\ion{H}{1} column density is \emph{larger} than inferred from
\ion{Si}{2} above. To this can be added the further complication
stemming from the merger event that the galaxy is currently undergoing,
which may have mixed gas of different metal contents. In any case,
though, we would expect the regions nearest to the starbursts to
generally have higher metallicity than the surrounding cool gas, such
that the HI column density inferred above is more likely to be
underestimated than overestimated.

@@

Looking back at Fig.~\ref{fig:SingleLines}, the profile of \ion{C}{2}
$\lambda 1334$ is clearly deeper than \ion{Si}{2} $\lambda 1260$, which
we had otherwise concluded is optically thick and thus provides a limit
for the velocity binned covering fractions. We believe the explanation
is a contribution from \ion{C}{2} $\lambda$ 1335.7 blending with
$\lambda 1334$. This transition has an oscillator strength of
$f_{ik} = 0.114$, compared to $f_{ik} = 0.127$ for $\lambda 1334$ (there
is a third line at 1335 Å, but that is an order of magnitude weaker), so
given a sufficient population of its ground level, it is strong enough
to make a non-negligible contribution to the resulting line profile. One
should however bear in mind that these lines rise from different lower
fine structure levels; $\lambda 1334$ arises from ²P$_{1/2}$, while the
other two arise from ²P$_{3/2}$. The population of the latter level
depends sensitively on physical conditions in its system of origin,
especially those in photo dissociation regions.

>>>>

\begin{abstract} \textbf{ Star forming galaxies are believed to be a major
		source of Lyman Continuum (LyC) radiation responsible for
		reionizing the early Universe. Direct observations of escaping
		ionizing radiation have however been few and with low escape
		fractions. In the local Universe, only ~10 emitters have been
		observed, with typical escape fractions of a few percent. The
		mechanisms regulating this escape need to be strongly evolving
		with redshift in order to account for the Epoch of Reionization.
		Gas content and star formation feedback are among the main
		suspects, known to both regulate neutral gas coverage and evolve
		with cosmic time. In this paper, we reanalyze HST-COS
		spectrocopy of the first detected local LyC leaker, Haro 11. We
		examine the connection between LyC leakage and Lyman-$\alpha$
		line shape, and feedback-influenced neutral ISM properties like
		kinematics and gas distribution. We discuss the two extremes of
		an optically thin, density bounded ISM and a riddled, optically
		thick, ionization bounded ISM, and how Haro 11 fits into their
		theoretical predictions.  We find that the most likely ISM model
		is a clumpy neutral medium embedded in a highly ionized medium
		with a combined covering fraction of unity and a residual
		neutral gas column density in the ionized medium high enough to
		be optically thick to Lyman-$\alpha$, but low enough to be at
		least partly transparent to Lyman continuum and undetected in Si
		II. This suggests that SF feedback and galaxy-scale interaction
		events play a major role in opening passageways for ionizing
radiation through the neutral medium.} \end{abstract}

%%% SUBMITTED ORIGINAL VERSION
% Star forming galaxies are believed to be a major source of ionizing
% radiation responsible for reionizing the early Universe. Direct
% observations of escaping ionizing radiation have however been few and
% with escape fractions far below what would be necessary to account for
% the ionization history of the Universe. In the local Universe, only a
% few handfuls of emitters have been observed with typical escape
% fractions of a few percent. It seems the mechanisms regulating this
% escape need to be strongly evolving with redshift, if galaxies are to be
% held responsible for reionization. Gas mass, star formation feedback and
% thus star formation activity are among the main suspect parameters,
% known to both regulate neutral gas coverage and evolve with cosmic time.
% In this paper, we present a reanalysis of far-UV HST-COS spectra of the
% first detected Lyman continuum leaker in the local Universe, Haro 11, in
% which we examine the connections between Lyman continuum leakage and
% Lyman-$\alpha$ line shape and feedback-influenced neutral ISM properties
% like kinematics, geometry and column density. We discuss the two
% extremes of an optically thin, density bounded ISM and a riddled,
% optically thick, ionization bounded ISM and how Haro 11 fits into the
% theoretical predictions made from these models. We find that the most
% likely ISM model for Haro 11 knot C is one of clumpy neutral medium
% embedded in a highly ionized inter-clump medium with a combined covering
% fraction of unity and a residual neutral gas column density in the
% ionized medium of
% $5 \times 10^{13} \leq N_{\rm HI} \leq 4 \times 10^{15}$ cm⁻²; high
% enough to be optically thick to Lyman-$\alpha$, but low enough to be at
% least partly transparent to Lyman continuum and undetected in
% \ion{Si}{2}.

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009, Hayes2009}, \textbf{since the resolution of absorption
lines is determined by the resolution of the background continuum source}. We 
then extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We modelled
vignetting in the COS aperture by multiplying each pixel in the circular
UV-continuum image with an interpolation of the values given in the
throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the aperture
cutouts were collapsed along the cross-dispersion direction, and the
FWHM of the flux distribution along the dispersion direction is reported
as the effective resolution of this observation. This is converted from
arc seconds to COS resolution elements by the factor 0.171 arcsec/resel
reported in the COS instrument handbook \citep{CosHandbook}

\begin{figure}
\centering
\includegraphics[width=1.000\hsize]{./EffResol.pdf}
\caption{Effective resolution estimate. \textbf{Upper left}: Haro 11 in
UV-continuum, data from \citet{Ostlin2009, Hayes2009}. Data is
shown on a square root scale and cut levels set for best detail. Inset
in blue is the COS aperture covering the region surrounding knot
B in the terminology of \citet{Vader1993}. N is up, E is left.
\textbf{Upper right}: The extracted, throughput-corrected in-aperture
image, rotated so horizontal corresponds to the dispersion direction of
the spectrograph. \textbf{Lower panel}: The collapsed flux profile in
the aperture, with the shaded region showing the width of the effective
resolution element, which is also given in arc seconds and resolution
elements/bins.}\label{fig:resol}
\end{figure}

The procedure is illustrated in Fig.~\ref{fig:resol}. Here, the upper left panel
shows Haro 11 in UV-continuum using HST imaging data from \citet{Ostlin2009};
\citet{Hayes2009}, with the COS aperture coverage from this observation overlaid
in blue. The upper right panel shows the galaxy as the COS saw it, with the
region covered by the aperture cut out, rotated and vignetting-corrected as
described above.  Finally, the lower panel shows the flux profile in the
aperture, collapsed along the cross dispersion direction. The FWHM of this
distribution has been adopted as the effective resolution for this observation.
We find $R_{\text{eff}} = 0.3 \arcsec = 1.75$ spectral resolution elements as
defined in the COS instrument handbook \cite{CosHandbook}, which at the
wavelength of observed Ly$\alpha$ corresponds to 31 km s⁻¹. \textbf{This value
might however be slightly underestimated, since the flux profile is not well 
described by a Gaussian profile, but has stronger wings due to the morphology of 
the target in the aperture, meaning that a larger fraction of the total flux is 
outside the FWHM.}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{./HISLISProfiles.pdf}
\caption{The \ion{Si}{2} and other LIS (\textbf{upper}) and \ion{Si}{4}
(\textbf{lower}) profiles included in this study. Dashes denote data
points masked out due to foreground contamination. \textbf{Data in this figure has
been smoothed to better show the line shapes, but all computations have been
carried out on the unsmoothed data.} Note the
high-velocity component in both \ion{Si}{4} lines, extending out to
$\sim 850$~km~s$^{-1}$.}\label{fig:SingleLines}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=0.800\hsize]{./AOD-details-example.pdf}
\caption{Illustration of the Apparent Optical Depth computations. 
\textbf{Upper panel}: the three \ion{Si}{2} lines in this analysis shown 
together, with the zero-velocity bin marked with a vertical red line. 
\textbf{Lower panel}: From this bin, the three measured relative intensities are 
plotted against $f\lambda$ as black dots, with vertical error bars showing the
error spectrum from the COS pipeline. The best-fit
function $I/I_0 (f \lambda)$ as defined above is drawn in magenta. Also shown 
for illustration are the two extreme examples of a very high column density
and unchanged covering fraction, and of a very high covering fraction
but significantly lower column density.}\label{fig:AOD}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{./LyACoverfracs.pdf}
\caption{\textbf{Upper panel}: Ly$\alpha$ profile of Haro 11 C, in
approximate units of the surrounding continuum level. \textbf{The full line
indicates the measured values smoothed by a 5 px. flat kernel}; surrounding shading
encloses the $\pm 1 \sigma$ error band (mostly comparable in width to
the line width). \textbf{Middle panel}: Black steps show the averaged
LIS line profile, smoothed by a 5px kernel. Surrounding gray shading
denotes the $\pm 1 \sigma$ confidence band. \textbf{Lower panel}: Same
as middle panel, but for the \ion{Si}{4} transitions.
}\label{fig:HisLisLya}
\end{figure}

The picture is what we would expect from a LyC leaker: \textbf{The
neutral absorption features are weak, and the ionized features are
strong, revealing a   highly ionized medium
in front of the central cluster} . This is also in good agreement with the
findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the
\ion{Si}{2} and \ion{Si}{4} line profiles, indicating that they likely
represent two different phases in the same higher density regions. These
regions will likely be ionized on the side facing the central cluster,
being photoionized by its hot, massive stars. This also implies that the
nebular emission outflow found in Rivera-Thorsen et al, submitted,
at least partially traces the same gas as the \ion{Si}{4} absorption
here and, by extension, also the neutral medium. Looking at
fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found
at around $v \approx -50$ km s⁻¹\textbf{. Interestingly, a component was also
found in nebular emission at this velocity in the observations of 
Rivera-Thorsen et al., submitted}. The outflow velocity
$v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully consistent with the
findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium
resonance absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an
overall, weak \emph{redshift} of $v = 32$ km s⁻¹. From
e.g.~fig.~\ref{fig:HisLisLya}, it is evident that while neutral gas
\emph{is} present at these velocities, the velocity is at odds with the
integrated velocity of $v = -149$ km s⁻¹ found in this work. We note
that \ion{Na}{1} has a very low ionization potential of only $\sim 5$
eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.
This is in agreement with the finding of \citet{Sandberg2013} that NaD,
despite being a strong transition, shows absorption of only $\sim 95\%$
of continuum level and is mostly found in small, optically thick clouds.
This suggests that the NaD is tracing only the densest and/or dustiest
regions which are more slowly accelerated by star formation feedback
than the surrounding, more dilute medium. In fig.~\ref{fig:apert}, dusty
regions in the aperture are apparent E and W of knot C. \textbf{Seen in the
better resolution of fig. 1 in \citet{Adamo2010}, these regions seem
like they might be connected by a narrow dust lane partially covering
the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.}

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of \ion{H}{1} around line center. We find a
Lyman $\alpha$ peak velocity of
$v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$ km s⁻¹ \textbf{relative to} the 
systemic velocity found by \citet{Sandberg2013}, derived from nebular emission
lines in the region around knot C. This velocity, as it is also
discussed in \citet{Verhamme2015}, is just consistent with their
theoretical predictions for a density-bounded, low-column density
system, albeit on the upper limit of their allowed range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in \textbf{the same way as in
Puschnig et al. (submitted to MNRAS) as follows}. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$ cm⁻²
for just one velocity bin. Since \ion{H}{1} gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario.

Furthermore, while Ly$\alpha$ radiative transfer is dominated by gas of
$v \sim v_0$, LyC is sensitive to \ion{H}{1} at \emph{all}
velocities. The total column density of \ion{Si}{2} depends on the
configuration of the neutral clouds. Assuming that the area covered by
gas in each velocity bin is completely randomly located, a lower limit
to the total column density is:
%
\begin{equation}
\label{NHImean}
N_{\rm Si II}^{\rm mean} = \sum_i N_{\text{HI}, i} f_{C, i}
\end{equation}
%
Summing this over $-450$ km s⁻¹ $< v < 150$ km s⁻¹, the velocity range
over which the column densities can be reasonably well determined (and
removing the unphysically high value in the bin at $v \approx 222$ km
s⁻¹), yields a lower limit of $\log N_{\rm Si II} = 13.52 \pm 0.15$ and
$\log N_{\rm H I} 19.01 \pm 0.17$, corresponding to $\approx 64$ optical
depths in LyC, strongly incompatible with an optically thin,
density-bounded scenario.

However, the \textbf{strong} riddled ionization bounded scenario 
\textbf{-- i.e. consisting of only a fully neutral and a fully ionized phase --} 
is easily
ruled out since the Ly$\alpha$ profile does not have any appreciable
emission component at zero velocity. We therefore expect a residual
neutral fraction to remain in the ionized phase; a fraction which has a
column density high enough to block Ly$\alpha$ efficiently at line
center, but low enough to be at least part transparent to LyC
radiation and undetected in \ion{Si}{2}.

We can estimate the lowest detectable $N_{\rm Si II}$ by noting that the
relative errors for the \ion{Si}{2} lines in our spectrum are
$\sim 0.05$. Assuming a covering fraction of unity for the dilute
neutral component, and adopting the oscillator strength of the strongest
of the \ion{Si}{2} transitions, $f\lambda_{1260 \AA} = 1486.8$, we find
in the limit that $I/I_0 = e^{-\tau} = 0.95$, and eq.~\ref{eq:tau}
becomes:
%
\begin{equation}
\label{eq:silim}
N_{\rm Si II}^{\rm min} = -\log_{e}(0.95) \frac{3.768 \times 10^{14}}{f\lambda} 
    = 10^{10.1} \text{cm}^{-2},
\end{equation}
%
 which with the adopted metallicity for Haro 11 C corresponds to a
minimum Hydrogen column density of
$N_{\rm HI}^{\rm min} \sim 4.0 \times 10^{15}$ cm⁻². This leaves around
2 orders of magnitude in $N_{\rm H I}$, in which the gas is not detected
in \ion{Si}{2} and is optically thick to Ly$\alpha$ while translucent to
LyC. If we require the gas to be detected in at least two of
the lines included in this analysis, the limiting hydrogen column
density becomes $N_{\rm HI}^{\rm min} \sim 4.9 \times 10^{16}$ cm⁻²,
adding another order of magnitude to the allowed range, but we adopt the
lower value as a conservative estimate.

The existence of a diffuse neutral component being present in the
ionized medium seems consistent with what is found in \textbf{Lyman Break
Analogs and Green Pea galaxies \citep{Heckman2011, Henry2015}}. 
These galaxies have sometimes very low
LIS absorption depths, which is usually indicative of low covering
fractions and high porosity of the neutral medium - and yet they find
covering fractions of near unity of \ion{H}{1} from Lyman $\beta$
absorption, indicating column densities of $N \gtrsim 10^{16}$ cm$^{-2}$. Like in
the case of Haro 11, this indicates that a non-negligible neutral
component must be present in the ionized phase, although it may very
well be of such low densities and metallicity that metal absorption from
this medium is undetectable.

The column densities we derive for \ion{Si}{2} in this work are
generally between $12.0 \lesssim \log N \lesssim 12.5$, column densities
around which the transitions involved become optically thick: For
$\lambda \lambda 1260, 1526 \text{ and } 1304$, $\tau$ becomes 1 at
$\log N \sim 11.3, 12.2$ and $12.5$, respectively. Off the regions of
strongest absorption, and in particular at $v_0$, only $\lambda 1260$
seems to be optically thick, and it seems the column densities arrived
at can be trusted. But at line center, around $v\sim v_{\mathrm{int}}$, the
column densities found are so close to the limit that they are most
likely to be interpreted as lower limits. There might also be
systematics in the determination of the continuum around the lines which
may lead to the lines at $\lambda \lambda 1304, 1526$ to be falsely seen
as shallower than $\lambda 1260$, meaning that all computed column
densities are really lower limits rather than actual values. The
confidence intervals in fig. \ref{fig:WithColDens} do \emph{not} reflect
this possible saturation, but only the formal errors from the best
approximation to the residual intensity, they do not include
systematics.

@@

The exact value of the metallicity is however uncertain. The values
found from nebular recombination lines by \citet{Guseva2012} are
measured mainly in the central \ion{H}{2} regions around the clusters;
and differ by 0.2 dex between knot B and C. The neutral, outflowing gas
could be mixed, or have an unseen LOS distance component larger than the
knot separation, drawing into question which is the better value to
assume for this gas. We base our conclusions on the value found for knot
C, but note that using the metallicity of $12 + \log (O/H) = 8.3$ found
for knot B by \citet{Guseva2012}, the inferred \ion{H}{1} column
densities are $3.9 \pm 0.6 \times 10^{17}$ cm$^{-2}$.
However, the question of the exact metallicity of the outflowing gas is
complicated. Gas closer to the star forming regions is expected to be
more strongly enriched than gas further away, which would imply that the
\ion{H}{1} column density is \emph{larger} than inferred from
\ion{Si}{2} above. To this can be added the further complication
stemming from the merger event that the galaxy is currently undergoing,
which may have mixed gas of different metal contents. In any case,
though, we would expect the regions nearest to the starbursts to
generally have higher metallicity than the surrounding cool gas, such
that the HI column density inferred above is more likely to be
underestimated than overestimated.

@@

Looking back at Fig.~\ref{fig:SingleLines}, the profile of \ion{C}{2}
$\lambda 1334$ is clearly deeper than \ion{Si}{2} $\lambda 1260$, which
we had otherwise concluded is optically thick and thus provides a limit
for the velocity binned covering fractions. We believe the explanation
is a contribution from \ion{C}{2} $\lambda$ 1335.7 blending with
$\lambda 1334$. This transition has an oscillator strength of
$f_{ik} = 0.114$, compared to $f_{ik} = 0.127$ for $\lambda 1334$ (there
is a third line at 1335 Å, but that is an order of magnitude weaker), so
given a sufficient population of its ground level, it is strong enough
to make a non-negligible contribution to the resulting line profile. One
should however bear in mind that these lines arise from different lower
fine structure levels; $\lambda 1334$ arises from ²P$_{1/2}$, while the
other two arise from ²P$_{3/2}$. The population of the latter level
depends sensitively on physical conditions in its system of origin,
especially those in photo dissociation regions.
